Assessing variations in the chemistry of fossil- and modern-*Pinus* pollen

### 2021-06-09 manuscript.Rmd
merging results and discussion; before disc parts integration; removed old pca/lda text

<!-- Put this into a table -->
<!-- Tiefer See formed during the last glacial period, while the Meerfelder Maar record is from the largest maar of the Westeifel Volcaniv field and formed ~35000 BP. -->
<!-- Tiefer See and Meerfelder Maar are varved cores and were the samples used in this study span two different time-periods. -->
<!-- The sub samples from Tiefer see are from the last 80 yrs of the core (1933 - 2011), while the samples from Meerfelder Maar cover the timeperiod from the late Allerød to the Younger Dryas/Holocene transition (~12900 - ~11000 BP).   -->

@@

```{r message=FALSE, warning=FALSE, include=FALSE}
pca_acet_select = readRDS(here("data","output","pca_plots.rds"))
lda_tables = readRDS(here("data","output","lda_tables.rds"))
```

# Results

```{r meanspectra, echo=FALSE, fig.cap="FRE", message=FALSE, warning=FALSE, out.width="100%"}
knitr::include_graphics(here("figures", "mean_plots.png"))
```

@@

<!-- 
Tiefer See pollen is separated from the other two cores, which cluster closer together.
This is the same for acetolysed and SPT pollen. 
A PCA-LDA model to differentiate fossil pollen into the three cores achieved ~60% accuracy to classify MFM and DAL pollen and 85% accuracy to identify TSK pollen (REF TO SM).

A peak at 1030 and ~1200 cm^-1^ is visible in the acetolysed pollen but not in the modern spectra or SPT pollen.
-->

@@

The 1090 cm^-1^ peak dissappears after acetolysis from all spectra as does the 1630 cm^-1^ peak in Tiefer See. 
The region 1000 - 1200 cm^-1^ is now dominated in acetolysed fossil pollen by two peaks and its shoulders at 1040 and 1170 cm^-1^.
This 1040 cm^-1^ has high importance for the PLSR model differentiating acetolysed and non acetolysed pollen (Fig. \@ref(fig:ordinations) e and f).
Using this PLSR model the acetolysed pollen is clearly separated from the non acetolysed pollen, with a much higher distance between samples for acetolysed pollen compared to non acetolysed pollen (Fig. \@ref(fig:ordinations) d).
In addition, there is some differentiation between the cores, samples from Tiefer See are more separate from samples of the other two cores, which show more similar scores in the ordination.

<!--
LDA to classify the pollen into depths showed higher accuracy for acetolysed pollen than SPT pollen (Table \@ref(tab:ldadepth)).
For the SPT pollen the accuracy increased when the depths were restricted to the same depths as the acetolysed samples.
Generally lower number of depths to classify increased accuracy.
sporopollenin peaks are similar between depths, changes in the other parts of the spectra are driving separation.
-->

>>>>

@@

# Results and Discussion

```{r meanspectra, echo=FALSE, fig.cap="FRE", message=FALSE, warning=FALSE, out.width="90%"}
knitr::include_graphics(here("figures", "mean_plots.png"))
```

## Fresh vs fossil pollen

@@

## Differences between fossil pollen
Using this PLSR model the acetolysed pollen is clearly separated from the non acetolysed pollen, with a much higher distance between samples for acetolysed pollen compared to non acetolysed pollen (Fig. \@ref(fig:ordinations) d).
In addition, there is some differentiation between the cores, samples from Tiefer See are more separate from samples of the other two cores, which show more similar scores in the ordination.

## Anomalies in fossil pollen spectra

@@

The 1090 cm^-1^ peak dissappears after acetolysis from all spectra as does the 1630 cm^-1^ peak in Tiefer See. 
The region 1000 - 1200 cm^-1^ is now dominated in acetolysed fossil pollen by two peaks and its shoulders at 1040 and 1170 cm^-1^.
This 1040 cm^-1^ has high importance for the PLSR model differentiating acetolysed and non acetolysed pollen (Fig. \@ref(fig:ordinations) e and f).

## Application for Palaeoecology

# Conclusions
